\LARGE{\bf{Natural variation in teosinte at the domestication locus \emph{teosinte branched1} (\emph{tb1})}}\footnotemark

### 2014-07-09 ResponseToReview.tex

Specific suggestions follow:

7) Introduction:
Page 4, Lines 7 - 9: You say "Additionally, many domesticates show reduced genetic diversity when compared to their wild progenitors, and an understanding of the distribution of diversity in the wild and its phenotypic eﬀects has become increasingly useful to crop improvement (Kovach and McCouch, 2008).".  Can you tell us why an understanding is useful? Can you give some examples?\\

\res{We have expanded on this concept in the introduction using \emph{Oryza sativa} as an example: "For example, \emph{Oryza rufipogon}, the wild progenitor of domesticated rice, has proven useful for the integration of a number of beneficial QTL controlling traits such as grain size and flowering time into domesticated rice (Kovach and McCouch 2008).}

8) Page 5, line 12 - 14: You say "The tb1 locus appears to play an important role in the shade avoidance pathway in Zea mays and other grasses and may therefore be crucial to the ecology of teosinte (Kebrom and Brutnell, 2007; Lukens and Doebley, 1999)." Please clarify; do the references belong to the first half of the sentence, supporting tb1 playing a role in Zea mays, or the second half of the sentence, supporting tb1 possibly being crucial to the ecology of teosinte? Which of the two distinct things have actually been studied? If the first, could you tell us how tb1 has been shown to work in maize under shade conditions? If the second, was it actually studied, or merely hypothesized in teosinte?

\res{We have clarified issues regarding the references for this section of the paper.}

9) Page 5, Lines 15 - 16: you say you "aim to characterize the distribution of the Hopscotch insertion in parviglumis, mexicana, and landrace maize, and to examine the phenotypic eﬀects of the insertion in parviglumis."  The mexicana seem a bit tacked on and not much discussed. Did you phenotype mexicana? Can you present the data here, if so? Does the analysis of this group of teosintes add anything to your conclusions?\\

\res{Reviewer 2 is correct in noting that ssp. \emph{mexicana} is not as prevalent in our paper as ssp. \emph{parviglumis}. This fact is due to limitations in access to accessions of ssp. \emph{mexicana} as field collections for ssp. \emph{mexicana} are not nearly as extensive as those of ssp. \emph{parviglumis}. Additionally ssp. \emph{mexicana} are more difficult to cultivate in a greenhouse setting. We felt that it would be important to include ssp. \emph{mexicana} in our study, especially since it is known that gene flow occurs between ssp. \emph{mexicana} and landrace maize. We have noted in our methods that we were only able to phenotype a population from which we had a large number of seed collections.}

10) Materials and Methods:
Page 5, lines 22 - 23: how many individuals per accession? A fixed number, or variable, and depending on what? This can be mentioned so the reader does not have to go to the appendix or supplements.

### 2014-07-09 manuscript.tex
edits
reviewer 2

In teosinte and other plants that grow at high population density, individuals detect competition from neighbors via the ratio of red to far-red light.  An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering \citep{KebromBrutnell2007}. The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} and other grasses and may therefore be crucial to the ecology of teosinte \citep{KebromBrutnell2007, LukensDoebley1999}.  In this study we aim to characterize the distribution of the \emph{Hopscotch} insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}. We use a combination of PCR genotyping for the \emph{Hopscotch} element in our full panel and sequencing of two small regions upstream of \emph{tb1} in a subset of teosinte populations to explore patterns of genetic variation at this locus. Finally, we test for an association between the \emph{Hopscotch} element and tillering phenotypes in a population of \emph{parviglumis}.

>>>>

In teosinte and other plants that grow at high population density, individuals detect competition from neighbors via the ratio of red to far-red light.  An increase in far-red relative to red light accompanies shading and triggers the shade avoidance syndrome: a suite of physiological and morphological changes such as reduced tillering, increased plant height and early flowering \citep{KebromBrutnell2007}. The \emph{tb1} locus appears to play an important role in the shade avoidance pathway in \emph{Zea mays} \citep{LukensDoebley1999} and other grasses \citep{KebromBrutnell2007} and may therefore be crucial to the ecology of teosinte. In this study we aim to characterize the distribution of the \emph{Hopscotch} insertion in \emph{parviglumis}, \emph{mexicana}, and landrace maize, and to examine the phenotypic effects of the insertion in \emph{parviglumis}. We use a combination of PCR genotyping for the \emph{Hopscotch} element in our full panel and sequencing of two small regions upstream of \emph{tb1} in a subset of teosinte populations to explore patterns of genetic variation at this locus. Finally, we test for an association between the \emph{Hopscotch} element and tillering phenotypes in a population of \emph{parviglumis}.
